---
title: "Visualise billboard data"
output: html_notebook
---
 
#Loading the relevant packages
```{r}
library(tidyverse)
library(ggthemes)
library(extrafont)
loadfonts()
par(family = "CMU Serif")
```

#Reading the data
```{r}
billboard <- read_csv("billboard.csv")
```

#Finding the number of songs, both average and per year.
```{r}
billboard_per_year <- billboard %>%
  group_by(year) %>%
  summarise(
    number_of_songs = sum(count)
  ) 

billboard_average <- billboard_per_year %>%
  summarise(
    average_songs = mean(number_of_songs)
  )
```

#Creating the relevant column that describes the popularity of a song. 
#Position is multiplied by it's count (fraction) to ensure weighted popularity (some songs are listed multiple times as they are tagged with several genres)
```{r}
billboard_pop_column <- billboard %>%
  group_by(genre, year) %>%
  mutate(
    pos = 101 - pos,
    popularity = pos*count,
  )
```


#Summing the data per genre and per year. 
#Selecting the largest genres from that to clean the figure.
```{r}
billboard_plot <-  billboard_pop_column %>%
  summarize(
    popularity = sum(popularity)
  ) %>%
  left_join(billboard_per_year) %>%
  mutate(
    popularity = popularity / number_of_songs
  ) %>%
  filter(
    year >= 1960,
    genre != "jazz",
    genre != "religious",
    genre != "disco",
    genre != "classical/soundtrack",
    genre != "swing",
    genre != "reggae",
    genre != "blues",
    genre != "country",
    genre != "folk",
    genre != "alternative/indie"
  )
```


#Making a pretty plot.
```{r}
ggplot(billboard_plot) +
  geom_line(mapping = aes(x = year, y = popularity)) +
    facet_grid(genre ~ .) +
  theme_hc() +
  xlab("Year") + ylab("Popularity") + 
  ggtitle("Popularity of genres") + 
  theme(strip.text.y = element_text(angle = 0, family = "CMU Serif"), text = element_text(family = "CMU Serif")) +
#  scale_y_continuous(trans = "reverse") +
#  theme(
#   strip.background = element_rect(
#     color="black", aes(fill = rainbow(5)[genre]), size=1.5, linetype="solid" 
#     )) + 
  ggsave("billboard_genres.png")
  
```
#Done!
