---
title: "Combine wiki and billboard"
output: html_notebook
---

The packages and fonts necessary.
```{r}
library(tidyverse)
library(ggthemes)
library(extrafont)
library(ggplot2)
loadfonts()
par(family = "CMU Serif")
```

Loading all the relevant data.
```{r}
billboard <- read_csv("billboard.csv")
musicians <- read_csv('Group_Project/genres_data.csv')
genres <- read_csv('Group_Project/genres-mapping.csv')
```


Adapt the wikipedia database to be plottable - this is the exact same as what happens in the wikipedia visualisation. The differences are: 'genre' and 'year' columns no longer have a capitalisation, the year is selected from 1980 onwards and only Pop music is selected.
```{r}
musicians <- left_join(musicians, genres, by='genre')

musicians <- musicians %>%
  mutate(
    many_genres = genre,
    genre = classified_genre,
    year = as.numeric(year)
  ) %>%
  select(-c(classified_genre)) %>%
  filter(year > 1980)

musicians_10 <- musicians %>%
  filter(genre == "Pop music")

Number_of_genres_10 <- musicians_10 %>%
  group_by(genre, year) %>%
  summarize(n_genres = n()) %>%
  arrange(desc(n_genres)) %>%
  select(-c(genre))
```

This is the exact same adaptation of the billboard dataset as in the billboard visualisation. The difference is that only the pop genre is selected.
```{r}
billboard_per_year <- billboard %>%
  group_by(year) %>%
  summarise(
    number_of_songs = sum(count)
  ) 

billboard_average <- billboard_per_year %>%
  summarise(
    average_songs = mean(number_of_songs)
  )

billboard_pop_column <- billboard %>%
  group_by(genre, year) %>%
  mutate(
    pos = 101 - pos,
    popularity = pos*count,
  )

billboard_plot <-  billboard_pop_column %>%
  summarize(
    popularity = sum(popularity)
  ) %>%
  left_join(billboard_per_year) %>%
  mutate(
    popularity = popularity / number_of_songs
  ) %>%
  filter(
    year >= 1980,
    genre == "pop"
  )
  
```

Plotting the graph! Including a separate scale on the right side.
```{r}
ggplot() +
  geom_line(data = Number_of_genres_10, mapping = aes(x = year, y = n_genres/10, color='Production')) +
  geom_line(data = billboard_plot, mapping = aes(x = year, y = popularity, color='Consumption')) +
  scale_y_continuous(sec.axis = sec_axis(~ . *10 , name = "New artists producing")) +
  theme_hc() + theme(text = element_text(family = "CMU Serif")) +
  scale_color_manual(values = c('red', 'blue')) +
  labs(
    y = "Inverted billboard position",
    x = "Year",
    colour = "Parameter" ) +
  ggtitle("Relative popularity and production of pop music") +
  ggsave("combined_plot.pdf")
```









