---
title: "Visualisation in R"
output: html_notebook
---

Installing all the packages we need. 
```{r}
#install.packages("ggthemes")
#install.packages("xtable")
#install.packages("extrafont")
#font_import(pattern = "lmroman*")
```

Loading the relevant packages.
```{r}
library(tidyverse)
library(xtable)
library(ggthemes)
library(RColorBrewer)
library(extrafont)
```

Make the right font.
```{r}
loadfonts()
par(family = "CMU Serif")
```

Loading the data.
```{r}
musicians <- read_csv('genres_data.csv')
genres <- read_csv('genres-mapping.csv')
```

Rewriting the genres to only include the 20 most common ones.
```{r}
musicians <- left_join(musicians, genres, by='genre')
```

Renaming the columns to "Genre" and "Year" for the relevant columns.
Defining a dataframe with the 10 most common genres out of the 20.
```{r}
musicians <- musicians %>%
  mutate(
    many_genres = genre,
    Genre = classified_genre,
    Year = as.numeric(year)
  ) %>%
  filter(year > 1900)

musicians_10 <- musicians %>%
  filter(Genre == "Pop music" |
         Genre == "Hip hop music" |
         Genre == "Rock music" | 
         Genre == "Alternative rock" | 
         Genre == "Pop rock" |
         Genre == "Country music" |
         Genre == "Folk music" |
         Genre == "Heavy metal music" |
         Genre == "Rhythm and blues" |
         Genre == "Soul music")
```

Defining two dataframes with the number of genres for each of the 20 and 10 most common genres. 
```{r}
Number_of_genres_10 <- musicians_10 %>%
  group_by(Genre, Year) %>%
  summarize(n_genres = n()) %>%
  arrange(desc(n_genres))

Number_of_genres_20 <- musicians %>%
  group_by(Genre, Year) %>%
  summarize(n_genres = n()) %>%
  arrange(desc(n_genres))
```

Plotting a count-graph with the 10 most common genres.
```{r}
ggplot(musicians_10) + 
  geom_count(mapping = aes(x = Year, y = Genre, color = Genre), alpha = 0.7) +
  guides(color = FALSE) + 
  scale_size_area() +
  theme_hc() + 
  theme(text=element_text(family="CMU Serif")) +
  ggtitle("New artists by genre") +
   ggsave("dotplot.pdf")
```

  
  
Plotting a linegraph with the 10 most common genres.
```{r}
ggplot(Number_of_genres_10) +
  geom_line(mapping = aes(x = Year, y = n_genres, color = Genre)) +
  xlab("Year") + ylab("Number of new artists") +
  ggtitle("New artists by genre") +
  theme_hc() +
  theme(text=element_text(family="CMU Serif")) +
  ggsave("Linegraph_artists.pdf") 
```

Plotting a linegraph with the 3 popular genres
```{r}
most_popular <- Number_of_genres_20 %>%
  filter(Genre == "Dance music" |
    Genre == "Pop  music" | 
    Genre == "Hip hop music") 

  ggplot(most_popular) +
  geom_line(mapping = aes(x = Year, y = n_genres, color = Genre)) +
  xlab("Year") + ylab("Number of new artists")
```

